---
title: "Stat 240: Lab 00"
subtitle: "Due January 20th 5:00 PM PST"
author: "Peter Khrissanov"
date: "January 16 2026"
format: html
---

This introductory lab is to ensure you have the computing environment
required for the rest of the semester set up. If you have taken
a previous class involving computing in R much of this will be
familiar. If not, this should still be straightforward.

<!-- This lab will be graded however you should submit it to crowdmark, -->
<!-- to ensure you are familiar with the submission process for future -->
<!-- problem sets. -->
This lab will be graded out of 1 point, to ensure you are familiar with 
the submission process for future problem sets.


# Problem 1

- To complete the lab you will need to ensure you have installed 
R and RStudio. See the installation instructions in the files section
of Canvas.

# Problem 2

- Next you should go to the `lab_00` folder in canvas and download
this file `lab_00.qmd`. __Save this file somewhere specific on your 
computer__. For example, create a folder for stat 240 and save it there.

We will talk a lot more about saving and organiizing files 
in future weeks, but this is just a very quick way to get started.
(If you know what an RStudio project is you can set one up for the class
before you do this. If not we will discuss it soon.)

Now open this file you have saved in RStudio.
You will see little grayed out chunks corresponding to code, for example.

```{r}
num_1 <- 101
num_2 <- 303

num_1 + num_2
```

If you press the green play button this code will run and show you the
result.
You can create these code chunks with the 
shortcut `Cmd/Ctrl + Option/Alt + i`.

## __To Do__. Create a code chunk and do a small simple calculation here.

```{r}
word1 <- 3
word2 <- 2
  
word1+word2
```





# Problem 3

Now we will see how to read in a simple data file into R, as seen
in the first class.
From canvas download `lab_00_data.csv` and __save it in the same folder
as the qmd file you previously downloaded__.

You should then be able to run the following code, which will read
this data into R and show you the first 6 rows of data.

```{r}
data <- read.csv("lab_00_data.csv")
head(data)
```


# Problem 4

Finally, we will create a very simple plot using this data, by running
the following code. You will need to have read the data in in the 
previous step for this code to run!

```{r}
plot(data$x, data$y)
```

# Problem 5

## To Do. Now create a pdf of your version of this document.

To do this you should:

- Modify the author section above, adding your name
- Render the quarto document, using the Render button at the top of
the document.

Then, this will `Render` the document,
creating a file with the same name but now with the `html` extension.
It will also open in the `Viewer` tab beside files, where you can see what
it looks like.

You can also open this `html` file by clicking on it, and opening it with your
web browser.

When you render a file, code written inside a code chunk will be run
and the output will be placed in the file created. This 
also means that if you write code which does not run 
properly, it will cause an error and RStudio will
not be able to knit your file.

<!-- # Saving Files for Submission -->

This is almost in the format we want to submit, except we need to
convert it to `pdf`. To do this, when it is open in the browser
you should `Print it to PDF`, saving it with the correct name.

This pdf is the file
you should then submit to crowdmark for both labs and computer based
exams in this course.
This lab is to ensure you have this set up correctly
for the rest of the semester!


